### 2020-02-13 stats/model_comparison.Rmd
update model sample processing file

---
title: "Model Comparison"
output: html_document
---

```{r}
library(tidyverse)
library(readxl)
library(matrixStats)
library(data.table)

estimate_mode <- function(s) {
  d <- density(s)
  return(d$x[which.max(d$y)])
}

hdi_upper<- function(s){
  m <- HPDinterval(mcmc(s))
  return(m["var1","upper"])
}

hdi_lower<- function(s){
  m <- HPDinterval(mcmc(s))
  return(m["var1","lower"])
}
```
# Data

```{r}
data_comb_pre <- read_csv("../data/combination.csv")

aoa_ratings_all <- read_xlsx(path = "../data/words_aoa_ratings.xlsx", sheet = 1)

aoa_ratings <- aoa_ratings_all %>%
  filter(Word %in% data_comb_pre$item )%>%
  mutate(mean_aoa = as.numeric(Rating.Mean),
         item = Word)%>%
  select(item,mean_aoa)

data_comb <- data_comb_pre%>%
  left_join(aoa_ratings)%>%
  ungroup()%>%
  mutate(item = fct_reorder(factor(item), mean_aoa), 
         source = "Data")

d_model_comp <- data_comb%>%
  select(-age)%>%
  mutate(age = round(age_month,2),
         item = as.character(item))%>%
  select(alignment, item, age, correct)

```

# Model comparison prediction

```{r}

d_model_comp.summarized <- d_model_comp %>% 
  group_by(alignment, item, age) %>% 
  summarize(n=n(), k = sum(correct))

# join predictions with the data

model_comp <- bind_rows(

  readRDS("../saves/model_pred_prag.rds") %>%
    right_join(d_model_comp.summarized) %>%
    select(model,chain,iteration, age, alignment,item,pred, n, k),

  readRDS("../saves/model_pred_global.rds") %>%
    right_join(d_model_comp.summarized)%>%
    select(model,chain,iteration, age, alignment,item,pred, n, k),

   readRDS("../saves/model_pred_flat.rds") %>%
    right_join(d_model_comp.summarized)%>%
    select(model,chain,iteration, age, alignment,item,pred, n, k),

  readRDS("../saves/model_pred_prior.rds") %>%
    right_join(d_model_comp.summarized)%>%
    select(model,chain,iteration, age, alignment,item,pred, n, k)
) %>%
  mutate(
    log_like = dbinom(x = k, size = n, prob = pred, log = TRUE)
  ) %>%
  group_by(chain, model, iteration) %>%
  summarize(total_likelihood_sample = sum(log_like)) %>%
  group_by(chain, model) %>%
  summarize(marginal_likelihood = matrixStats::logSumExp(total_likelihood_sample)) %>%
  spread(model, marginal_likelihood)%>%
  mutate(prag_vs_global = pragmatic - global, # subtract likelihood for each predictions and model
         prag_vs_flat= pragmatic - flat,
         prag_vs_prior= pragmatic - prior,
         global_vs_flat = global - flat,
         global_vs_prior = global - prior,
         flat_vs_prior = flat - prior) %>%
  select(-pragmatic, -global, -flat, - prior) %>%
  gather(model, logBF, -chain)

```

# Model comparison explanation

```{r}

d_model_comp.summarized <- d_model_comp %>% 
  group_by(alignment, item, age) %>% 
  summarize(n=n(), k = sum(correct))

model_like <- bind_rows(

  readRDS("../saves/model_pred_prag.rds") %>%
    right_join(d_model_comp.summarized) %>%
    select(model,chain,iteration, age, alignment,item,pred, n, k),

  readRDS("../saves/model_pred_prag_fb.rds") %>%
    mutate(model = "fullBayes")%>%
    right_join(d_model_comp.summarized)%>%
    select(model,chain,iteration, age, alignment,item,pred, n, k)

    # ,readRDS("../saves/model_pred_mm.rds") %>%
    # mutate(model = "mixture")%>%
    # right_join(d_model_comp.summarized)%>%
    # select(model,chain,iteration, age, alignment,item,pred, n, k)
) %>%
  mutate(
    log_like = dbinom(x = k, size = n, prob = pred, log = TRUE)
  ) %>%
  group_by(chain, model, iteration) %>%
  summarize(total_likelihood_sample = sum(log_like))

# variation across chains

model_by_chain <- model_like %>%
  na.omit(chain)%>%
  group_by(model, chain) %>%
  summarize(marginal_likelihood = matrixStats::logSumExp(total_likelihood_sample)) %>%
  spread(model, marginal_likelihood)

# compare models
model_comp <- model_like %>%
  na.omit(chain)%>%
  group_by(model) %>%
  summarize(marginal_likelihood = matrixStats::logSumExp(total_likelihood_sample)) %>%
  spread(model, marginal_likelihood)%>%
  mutate(fb_vs_free = fullBayes - pragmatic
         # ,fb_vs_mixture= fullBayes - mixture,
         # free_vs_mixture = pragmatic - mixture
         ) %>%
  select(-pragmatic, -fullBayes, -mixture) %>%
  gather(model, logBF)

```