### 2020-03-31 paper/visualizations.Rmd
work on MHs comments

---
title: "Word learning at the semantics pragmatics interface"
output: html_document
---

```{r, include = F}
library(tidyverse)
library(rwebppl)
library(gridExtra)
library(coda)
library(ggthemes)
library(readxl)
library(ggpubr)
library(stringr)
library(matrixStats)
library(data.table)
library(langcog)
library(ggimage)
library(BayesFactor)

library(cowplot)
library(magick)

estimate_mode <- function(s) {
  d <- density(s)
  return(d$x[which.max(d$y)])
}

hdi_upper<- function(s){
  m <- HPDinterval(mcmc(s))
  return(m["var1","upper"])
}

hdi_lower<- function(s){
  m <- HPDinterval(mcmc(s))
  return(m["var1","lower"])
}

knitr::opts_chunk$set(echo=F, warning=FALSE, message=FALSE, size="small")

```

```{r data}
aoa_ratings <- read_xlsx(path = "../data/words_aoa_ratings.xlsx", sheet = 1)%>%
  filter(Word %in% c("carrot","duck","bread","apple","kite","horseshoe","plug","garlic","barrel","eggplant","pawn","papaya"))%>%
  mutate(mean_aoa = as.numeric(Rating.Mean),
         item = Word)%>%
  select(item,mean_aoa)

me_data <- read_csv("../data/me.csv")
prior_data <- read_csv("../data/novelty.csv")
comb_data <- read_csv("../data/combination.csv")%>%
  mutate(model = "data")%>%
  left_join(aoa_ratings) %>%
  ungroup()%>%
  mutate(item = fct_reorder(factor(item), mean_aoa))

```

```{r}
# parameter distributions based on model 
item_params <- readRDS("../saves/item_params.rds")
global_params <- readRDS("../saves/global_params.rds")
item_sigma <- readRDS("../saves/item_sigma.rds")

# summaries
item_params_summary <- item_params %>%
  group_by(item, parameter)%>%
  summarise(mode = estimate_mode(value),
            uci = hdi_upper(value),
            lci = hdi_lower(value))

global_params_summary <- global_params %>%
  group_by (parameter,type)%>%
  summarise(mode = estimate_mode(value),
            uci = hdi_upper(value),
            lci = hdi_lower(value))

# selection object
select <- sample(1:length(unique(global_params$iteration)), 30)

# plot for semantic knowledge

sem_know_map <- item_params_summary%>%
  select(-uci,-lci)%>%
  spread(parameter, -item) %>%
  expand_grid(., age = unique(me_data$age)) %>%
  mutate(sem_know = plogis(intercept + slope * age)) %>%
  left_join(aoa_ratings) %>%
  ungroup()%>%
  mutate(item = fct_reorder(factor(item), mean_aoa))

plot_sem_know <- ggplot()+
  geom_line(data = sem_know_map, aes(x = age+2, y= sem_know, col = item, group = item),size = 1)+
  ylab("Semantic knowledge")+
  xlab("Age")+
  ylim(0,1)+
  theme_few()+
  scale_colour_viridis_d(name = "Object")

# plot or speaker optimality

speak_opt <- global_params%>%
  filter(parameter == "speaker_optimality")

speak_opt_line <- speak_opt%>%
  filter(iteration %in% select)%>%
  spread(type, value)%>%
  expand_grid(., age = unique(me_data$age))%>%
  mutate(y = intercept + slope * age)

speak_opt_map <- global_params_summary%>%
  ungroup()%>%
  filter(parameter == "speaker_optimality")%>%
  select(type, mode)%>%
  spread(type, mode)%>%
  expand_grid(., age = unique(me_data$age))%>%
  mutate(y = intercept + slope * age)

plot_speak_opt <- ggplot() +
  geom_line(data = speak_opt_line, aes(x=age+2,y = y, group = interaction(chain,iteration)), size = .2, alpha = .1)+
  geom_line(data = speak_opt_map, aes(x=age+2,y = y),size = 1)+
  labs(x="Age",y="Speaker informativeness")+
  theme_few() +
  ylim(0,10)+
    xlim(2,5)+
  guides(alpha = F, fill = F, col = F)

# Plot prior sensitivity

prior <- global_params%>%
  filter(parameter == "prior")

prior_line <- prior%>%
  filter(iteration %in% select)%>%
  spread(type, value)%>%
  expand_grid(., age = unique(me_data$age))%>%
  mutate(y = plogis(intercept + slope * age))

prior_map <- global_params_summary%>%
  ungroup()%>%
  filter(parameter == "prior")%>%
  select(type, mode)%>%
  spread(type, mode)%>%
  expand_grid(., age = unique(me_data$age))%>%
  mutate(y = plogis(intercept + slope * age))

plot_prior <- ggplot() +
  geom_line(data = prior_line, aes(x=age+2,y = y, group = interaction(chain,iteration)), size = .2, alpha = .1)+
  geom_line(data = prior_map, aes(x=age+2,y = y),size = 1)+
  labs(x="Age",y="Prior sensitivity")+
  theme_few() +
  ylim(0,1)+
  xlim(2,5)+
  guides(alpha = F, fill = F, col = F)

```
```{r}
plot_cor_model_pred <- readRDS( "../saves/corr_model_data_pred.rds")%>%
  mutate(model = factor(model, levels = c("integration", "no word knowledge", "no common ground", "no mutual exclusivity")))%>%
  filter(model == "integration")

pred_cor <- ggplot(data = plot_cor_model_pred,aes(x = model_mean, y = data_mean, col = alignment)) +
  geom_abline(intercept = 0, slope = 1, lty = 2, alpha = 1, size = .5)+
  geom_errorbar(aes(ymin = data_ci_lower, ymax = data_ci_upper),width = 0,size = .5, alpha = .7)+
  geom_errorbarh(aes(xmin = model_ci_lower, xmax = model_ci_upper), height = 0,size = .5, alpha = .7)+
  geom_point(size = 1.5, stroke = 1, pch = 5)+
  #coord_fixed()+
 stat_cor(method = "pearson", label.x = 0.01, label.y = 0.99, aes(x = model_mean, y = data_mean), inherit.aes = F, size = 3)+
  xlim(0,1)+ylim(0,1)+
  xlab("Model")+
  ylab("Data")+
  facet_grid(~age)+
  theme_few() + 
  scale_colour_ptol(name ="Condition")
```

```{r}
plot_prag_pred <- readRDS( "../saves/plot_prag_pred.rds")

pred_pred <- ggplot(data = comb_data, aes(x = age_num, y = correct)) +
  #geom_hline(yintercept = 0.5, lty=2)+
  #geom_jitter(col = "black", height = .025, alpha = .1)+
  #geom_pointrange(data = comb_data_binned, aes(x = subage+.5, y = data_mean, ymin =data_ci_lower, ymax = data_ci_upper), stroke = 1, pch = 5)+
  #geom_line(data = comb_data_binned, aes(x = subage+.5, y = data_mean))+
  geom_line(data = plot_prag_pred, aes(x=age+2,y = pred, col = alignment, group = interaction(chain,iteration)), size = .2, alpha = .25)+
  geom_smooth(data = comb_data, aes(x = age_num, y = correct), col = "black", method = "glm", method.args = list(family = "binomial"), se = T, alpha = .4, lty = 2, size = 1)+
  labs(x="Age",y="Mutual Exclusivity effect")+
  facet_grid(alignment~item)+
  theme_few() +
  ylim(-0.05,1.05)+
  xlim(2,5)+
  guides(alpha = F, fill = F, col = F)+ 
  scale_colour_ptol(name = NULL)
```

```{r}

ggarrange(
ggarrange(plot_speak_opt, plot_prior,plot_sem_know, labels = c("A","B","C"), nrow = 1, widths = c(1,1,1.5)),
ggarrange(pred_cor, labels = c("D"), nrow = 1 ),
nrow= 2,
heights = c(1,1)
)

```

```{r}
ggsave("../graphs/paper_figure2.png", width = 5, height = 3.5, scale = 2)
```