### 2021-02-26 stats/model_comparison.Rmd
clean repo for revision

---
title: "Model Comparison"
output: html_document
---

```{r}
library(tidyverse)
library(matrixStats)
library(ggthemes)

estimate_mode <- function(s) {
  d <- density(s)
  return(d$x[which.max(d$y)])
}

hdi_upper<- function(s){
  m <- HPDinterval(mcmc(s))
  return(m["var1","upper"])
}

hdi_lower<- function(s){
  m <- HPDinterval(mcmc(s))
  return(m["var1","lower"])
}
```

# Model comparison

Model predictions can be downloaded via the following link: https://keeper.mpdl.mpg.de/d/7ae3805dba064bada8e5/
In the code below, model predictions are assumed to be in the sub-folder called `webppl/output/`.

```{r}

# sample <- bind_rows(
#   read_csv("../webppl/output/llh_model_chain1.csv")%>%mutate(chain = 1),
#   read_csv("../webppl/output/llh_model_chain2.csv")%>%mutate(chain = 2),
#   read_csv("../webppl/output/llh_model_chain3.csv")%>%mutate(chain = 3),
#   read_csv("../webppl/output/llh_model_chain7.csv")%>%mutate(chain = 4),
#   read_csv("../webppl/output/llh_model_chain8.csv")%>%mutate(chain = 5),
#   read_csv("../webppl/output/llh_model_chain9.csv")%>%mutate(chain = 6),
#   read_csv("../webppl/output/llh_model_chain10.csv")%>%mutate(chain = 7),
#   read_csv("../webppl/output/llh_model_chain11.csv")%>%mutate(chain = 8),
#   read_csv("../webppl/output/llh_model_chain12.csv")%>%mutate(chain = 9),
#   read_csv("../webppl/output/llh_model_chain13.csv")%>%mutate(chain = 10)
# )
# 
# sample_like <-  sample%>%
#   filter(a == "likelihood")%>%
#   rename(type = a,
#          model = b, 
#          alignment = c,
#          age = d,
#          item = e, 
#          log_like = f)%>%
#   mutate(age = as.numeric(age))%>%
#   group_by(chain, model, iteration) %>%
#   summarize(total_likelihood_sample = sum(log_like))
# 
# saveRDS(sample_like, "../saves/likelihood_sample.rds")

sample_like <- readRDS("../saves/likelihood_sample.rds")

# model_by_chain <- sample_like %>%
#   group_by(chain, model) %>%
#   summarize(marginal_likelihood = matrixStats::logSumExp(total_likelihood_sample)) %>%
#   spread(model, marginal_likelihood)
# 
# saveRDS(model_by_chain, "../saves/model_likelihood_chain.rds")

model_by_chain <-  readRDS("../saves/model_likelihood_chain.rds")

# model_like <- sample_like %>%
#   group_by(model) %>%
#   summarize(marginal_likelihood = matrixStats::logSumExp(total_likelihood_sample))
# 
# saveRDS(model_like, "../saves/model_likelihood.rds")

model_like <-  readRDS("../saves/model_likelihood.rds")

# model_comp <- model_like  %>%
#   group_by(model) %>%
#   spread(model, marginal_likelihood)%>%
#   mutate(prag_vs_global = pragmatic - global, # subtract likelihood for each predictions and model
#          prag_vs_flat= pragmatic - flat,
#          prag_vs_prior= pragmatic - prior_only,
#          prag_vs_bias = pragmatic - bias,
#          prag_vs_devBias = pragmatic - devBias,
#          devBias_vs_bias = devBias - bias,
#          global_vs_flat = global - flat,
#          global_vs_prior = global - prior_only,
#          flat_vs_prior = flat - prior_only) %>%
#   select(-pragmatic, -global, -flat, - prior_only, -bias, -devBias) %>%
#   gather(model, logBF)
# 
# saveRDS(model_comp, "../saves/model_comparison.rds")

model_comp <- readRDS( "../saves/model_comparison.rds")
```

# visualize log-likelihood distributions

```{r}
p_sample_like <- sample_like%>%
  mutate(model = fct_recode(model,
                            "integration" = "pragmatic",
                            "no word knowledge" = "global",
                            "no common ground" = "flat",
                            "no speaker informativeness" = "prior_only",
                            "biased" = "bias",
                            "developmental bias" = "devBias"),
         model = factor(model, levels = c("integration","no word knowledge","no common ground", "no speaker informativeness","biased", "developmental bias")))

p_model_like <- model_like %>%
  mutate(model = fct_recode(model,
                            "integration" = "pragmatic",
                            "no word knowledge" = "global",
                            "no common ground" = "flat",
                            "no speaker informativeness" = "prior_only",
                            "biased" = "bias",
                            "developmental bias" = "devBias"),
         model = factor(model, levels = c("integration","no word knowledge","no common ground", "no speaker informativeness","biased", "developmental bias")))

ggplot(p_sample_like, aes(x = total_likelihood_sample, col = model, fill = model))+
  geom_density(alpha = .5)+
  geom_vline(data = p_model_like, aes(xintercept = marginal_likelihood, col = model), lty = 2)+
  theme_minimal()+
  labs(x ="log likelihood (sample)", y = "")+
  scale_fill_colorblind(name = "Model")+
  scale_color_colorblind(name = "Model")
```